A multi-species benchmark for training and validating mass spectrometry proteomics machine learning models

### 2024-08-04 doc/nine-species/nine-species.tex
A few small updates

\title{A multi-species benchmark for training and validating large scale mass spectrometry proteomics machine learning models}

@@

We assigned peptide labels to spectra using the Tide search engine followed by post-processing with Percolator.
In creating the Tide index, we specified specified Cys carbamidomethylation as a static modification and allowed for the following variable modifications: Met oxidation, Asn deamidation, Gln deamidation, N-term acetylation, N-term carbamylation, N-term NH$_{3}$ loss, and the combination of N-term carbamylation and NH$_{3}$ loss by using the tide-index options {\tt --mods-spec 1M+15.994915,\allowbreak 1N+0.984016,\allowbreak 1Q+0.984016 --nterm-peptide-mods-spec 1X+42.010565,\allowbreak 1X+43.005814,\allowbreak 1X-17.026549,\allowbreak 1X+25.980265 --max-mods 3}.
Note that one of the nine experiments (\textit{Mus musculus}) was performed using SILAC labeling, but we searched without SILAC modifications and hence include in the benchmark only PSMs from unlabeled peptides.
Tide automatically added to each index a shuffled decoy peptide corresponding to each target peptide.
Thereafter, each MGF file was searched against the corresponding index using the precursor window size and fragment bin tolerance specified in the original study (Table~\ref{tab:benchmark}).
The search engine employed XCorr scoring with Tailor calibration \cite{sulimov2020tailor}, and we allowed for 1 isotope error in the selection of candidate peptides.
All search results were then analyzed jointly per species using the Crux implementation of Percolator, with default parameters.
For the benchmark, we retained all PSMs with Percolator q value $<$ 0.01.
We identified 13 MGF files with fewer than 100 accepted PSMs, and we eliminated all of these PSMs from the benchmark.
At this point in the processing pipeline, the dataset contains 2,898,611 annotated spectra (PSMs)  drawn from 339 RAW files and associated with 179,256 distinct peptides.

@@

The dataset contains files resulting from various steps in the generation of the benchmark:
\begin{itemize}
\item Spectrum files in MGF format, produced by msconvert.
\item Reference proteome files in FASTA format, downloaded from Uniprot.
\item Search results files for both targets and decoys, in tab-delimited format, produced by Tide.
\item PSM-level Percolator results files for targets, in tab-delimited format.
\item Annotated MGF files for both versions of the benchmark (main and balanced).
\end{itemize}
Also included are log files for the steps of the analysis pipeline carried out using Crux \cite{park2008rapid} (Tide indexing, Tide search, and Percolator).
\fixme{Add Zenodo URLs.}

@@

%\paragraph{Author Contributions}

>>>>

\title{A multi-species benchmark for training and validating mass spectrometry proteomics machine learning models}

@@

\begin{sloppypar}
We assigned peptide labels to spectra using the Tide search engine followed by post-processing with Percolator.
In creating the Tide index, we specified specified Cys carbamidomethylation as a static modification and allowed for the following variable modifications: Met oxidation, Asn deamidation, Gln deamidation, N-term acetylation, N-term carbamylation, N-term NH$_{3}$ loss, and the combination of N-term carbamylation and NH$_{3}$ loss by using the tide-index options {\tt --mods-spec 1M+15.994915,\allowbreak 1N+0.984016,\allowbreak 1Q+0.984016 --nterm-peptide-mods-spec 1X+42.010565,\allowbreak 1X+43.005814,\allowbreak 1X-17.026549,\allowbreak 1X+25.980265 --max-mods 3}.
Note that one of the nine experiments (\textit{Mus musculus}) was performed using SILAC labeling, but we searched without SILAC modifications and hence include in the benchmark only PSMs from unlabeled peptides.
Tide automatically added to each index a shuffled decoy peptide corresponding to each target peptide.
Thereafter, each MGF file was searched against the corresponding index using the precursor window size and fragment bin tolerance specified in the original study (Table~\ref{tab:benchmark}).
The search engine employed XCorr scoring with Tailor calibration \cite{sulimov2020tailor}, and we allowed for 1 isotope error in the selection of candidate peptides.
All search results were then analyzed jointly per species using the Crux implementation of Percolator, with default parameters.
For the benchmark, we retained all PSMs with Percolator q value $<$ 0.01.
We identified 13 MGF files with fewer than 100 accepted PSMs, and we eliminated all of these PSMs from the benchmark.
At this point in the processing pipeline, the dataset contains 2,898,611 annotated spectra (PSMs)  drawn from 339 RAW files and associated with 179,256 distinct peptides.
\end{sloppypar}

@@

The dataset contains files resulting from various steps in the generation of the benchmark:
\begin{itemize}
\item Spectrum files in MGF format, produced by msconvert.
\item Reference proteome files in FASTA format, downloaded from Uniprot.
\item Search results files for both targets and decoys, in tab-delimited format, produced by Tide.
\item PSM-level Percolator results files for targets, in tab-delimited format.
\item Annotated MGF files for both versions of the benchmark (main and balanced).
\end{itemize}
Also included are log files for the steps of the analysis pipeline carried out using Crux \cite{park2008rapid} (Tide indexing, Tide search, and Percolator).
The data is available at \url{https://doi.org/10.5281/zenodo.12819175}.

@@

\paragraph{Author Contributions}
WSN prepared the dataset, performed the validation experiments, and wrote the manuscript.
